Supplementary Data

### 2021-11-17 products/manuscript/Manuscript.Rmd
Miscellaneous_edits - CC
Intro & Supplementary File - CC

@@

@@

_When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates). However, these categories are usually analyzed separately to assess stream health. As such, my question is how well are these parameters correlated with each other. Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?_

_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories. The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._

@@

_There is 3,075 rows and 12 columns... So, a lot of data. This is because data has been collected for ~20 years. Unfortunately, I am seeing a lot of NA's in the data. Because this is citizen science monitoring data, NA most likely means the data wasn't collected during that event and is a form of human error bias._

_4. Use NA assessment to see determine which variables to remove from assessment. This was done using the is_na function._
_Several of the columns have more than half of the data missing. This is pretty common for citizen science data, seeing as collection is opportunistic. As such, I decided to remove P04 and temperature, because they are missing a lot of data. Finally, I decided to remove all missing values and see the condition of the data knowing that there are now no missing values. I removed NA data using the na.omit() function_

>>>>

```{r, echo=FALSE, results='hide', message=FALSE}
library(here)
library(tidyverse)
```

@@

_Stream health and water quality is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates). However, these categories are usually analyzed separately to assess stream health [cite]. If the impact of urbanization on freshwater bodies is not monolithic, we could be spending a lot of time and effort on monitoring programs that are not effective. Studies have shown conflicting results on what variables like conductivity, turbidity, e. coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality [cite]. Therefore, a study was conducted assess how well all these parameters correlated with each other in the context of the Adopt-a-Stream method. Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?_

_Data was acquired from a publicly available spreadsheet of water quality measurements collected by the Upper Oconee Watershed Network (UOWN) between 2003 and 2020. Data originated from three distinct watersheds in the Athens- Clarke County area: Middle Oconee (MIDO), North Oconee (NORO), and Oconee (BIOC). The data were analyzed first by using bivariate linear modeling of the variables conductivity, turbidity, nitrate (no3) concentration, E. coli cfu, and biodiversity indicator score (biological score). E. coli cfu and biological score were used as outcomes to specifically see if easy to measure physical/chemical properties could predict biological variables to assess water quality. Second, the MIDO data were analyzed using LASSO linear modeling to construct the best combinations of variables to predict E. coli cfu and biological score._

_Results indicated that the only significant bivariate relationships that existed were between E. coli cfu and conductivity for the MIDO site only. LASSO modeling produced models that did not fit the data any better than the null model that was calculated for cross validation. This indicates that physical/chemical water quality measurements cannot predict biological measurements. This suggests that the citizen science approach to water quality monitoring should be re-evaluated for sampling robustness or that the effect of urbanization is spatially and temporally different and may not captured well without more in-depth sampling. Future studies should investigate how specific aspects of urbanization actually impact our indicator variables of water quality and stream health._

@@

_The Earth's surface contains a relative proportion of approximately 3% freshwater to 97% salt water. As such, that three percent is supposedly meant to sustain a growing global population of 9 billion people. In a world where water scarcity is more and more common, our freshwater reserves are also threatened by pollution by urbanization. Urbanization surrounding a freshwater body creates a watershed dominated by impervious surface in the form of concrete and asphalt [cite]. Studies have shown that there is a significant negative relationship between impervious surface coverage and water quality. This is most likely due to accumulation and concentration of run-off pollutants from an entire watershed into a single drainage basin (i.e. freshwater body)[cite]. Because the urban world continues to grow with population, it is likely that we will see a proportional decrease in water quality in our freshwater bodies as a result. Therefore, a rapid and consistent monitoring system needs to be in place to effectively allocate resources to remediating body of water in the midst of being polluted._

_Currently, stream health and water quality is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates). Typically, these measurements include conductivity (measure of dissolved ions), dissolved oxygen, temperature, pH, nitrate and phosphorus concentration, turbidity (measure of suspended particulate matter in the water column), fecal indicator bacteria (measure of waste pollution), and benthic macroinvertebrate diversity indices (measure of impervious surface run-off on sensitive aquatic taxa) [cite]. However, these categories are usually analyzed separately to assess stream health [cite]. As such, it is hard to evaluate whether urbanization affects watersheds and water quality in fairly similar ways, regardless of location. Evaluating how uniform the impact of urbanization is on different watersheds would allow for a cookie-cutter monitoring and remediation method, which would allow implementation to be fairly easy for environmental managers. On the other hand, if the impact of urbanization on freshwater bodies is not monolithic, we could be spending a lot of time and effort on monitoring programs that are not effective. Currently, citizen science groups, like Adopt-a-Stream in Georgia, are collecting data on many watersheds throughout the world in an effort to consistently monitor water quality in a community. However, studies have shown conflicting results on what variables like conductivity, turbidity, e. coli cfu (fecal coliform indicator), nutrient (nitrogen and phosphorus) concentration, and biodiversity indicator score actually say about water quality [cite]. Therefore, my question is how well are these parameters correlated with each other in the context of the Adopt-a-Stream method. Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe the others to make a statement about overall stream health?_

_I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact to stream health in the other categories. The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find._

@@

_There is 3,075 rows and 12 columns. This is because data has been collected for ~20 years. Unfortunately, I am seeing a lot of NA's in the data. Because this is citizen science monitoring data, NA most likely means the data wasn't collected during that event and is a form of human error bias._

_4. Use NA assessment to see determine which variables to remove from assessment. This was done using the is_na function. Several of the columns have more than half of the data missing. This is pretty common for citizen science data, seeing as collection is opportunistic. As such, I decided to remove P04 and temperature, because they are missing a lot of data. Finally, I decided to remove all missing values and see the condition of the data knowing that there are now no missing values. I removed NA data using the na.omit() function_

### 2021-11-17 results/Supplementary_Data.Rmd
Intro & Supplementary File - CC

>>>>

_Figure 7: Scatter plot of LASSO model predicted Biological Score data as a function of actual Biological Score from the MIDO data set._
```{r}
lasso_prediction_plot_BS_location <- here::here("results", "lasso_prediction_plot_BS.png")

include_graphics(lasso_prediction_plot_BS_location)
```
</br>

_Figure 8: Scatter plot of LASSO model predicted E. coli cfu data as a function of actual E. coli cfu from the MIDO data set._
```{r}
lasso_prediction_plot_EC_location <- here::here("results", "lasso_prediction_plot_EC.png")

include_graphics(lasso_prediction_plot_EC_location)
```

```{r}
MIDO_ecfu_con_fit_location <- here::here("results", "MIDO_ecfu_con_fit.png")

include_graphics(MIDO_ecfu_con_fit_location)
```
</br>
_Table 1_

```{r}
MIDO_ecfu_con_RMSE_location <- here::here("results", "MIDO_ecfu_con_RMSE.png")

include_graphics(MIDO_ecfu_con_RMSE_location)
```
</br>
_Table 2_

```{r}
Null_bs_RMSE_location <- here::here("results", "Null_bs_RMSE.png")

include_graphics(Null_bs_RMSE_location)
```
</br>
_Table 3_

```{r}
Null_ecfu_RMSE_location <- here::here("results", "Null_ecfu_RMSE.png")

include_graphics(Null_ecfu_RMSE_location)
```
</br>
_Table 4_

```{r}
MIDO_bs_LASSO_top5_location <- here::here("results", "MIDO_bs_LASSO_top5.png")

include_graphics(MIDO_bs_LASSO_top5_location)
```
</br>
_Table 5_

```{r}
MIDO_ecfu_LASSO_top5_location <- here::here("results", "MIDO_ecfu_LASSO_top5.png")

include_graphics(MIDO_ecfu_LASSO_top5_location)
```
</br>
_Table 6_
